---
title: "EMUs Stage 3 Recruitment progress"
format: 
  closeread-html:
    theme: cosmo
    cr-style:
      narrative-background-color-sidebar: "#f55845"
      section-background-color: white
      narrative-text-color-sidebar: black
      narrative-font-size: 2em
editor: visual
execute:
  echo: false
  warning: false
editor_options: 
  chunk_output_type: console
theme: minty
knitr:
  opts_chunk:
    out.width: "8in"
---

```{r}
library(tidyverse)
library(scales)
library(tidymodels)
theme_set(theme_classic(base_size = 16))
library(pins)
board_local = board_folder("/home/common/emus/")
emus_patients = board_local %>% pin_read("emus_recruitment")

emus_colours = c("#d73027", "#fdae61", "#66bd63")
```

```{r}
weekly_recruitment = emus_patients %>% 
  count(week) %>% 
  mutate(total = cumsum(n))

study_dates = tibble(week = ymd("2024-08-05") + weeks(1:100)) %>% 
  filter(week < ymd("2026-01-01"))

# base R + broom
#lm(total~week, data = weekly_recruitment) %>% augment(newdata = future)

# tidymodels workflow for linear regression
recruitment_lm       = linear_reg()
recruitment_recipe   = recipe(total~week, data = weekly_recruitment)
recruitment_workflow = workflow() %>% 
  add_recipe(recruitment_recipe) %>% 
  add_model(recruitment_lm)

filter_sets = function(df, set = c("thisyear", "fourweeks")){
  if (set == "thisyear"){
    df %>% 
      filter(week > ymd("2024-12-31"))
  } else if(set == "fourweeks"){
    df %>% 
      filter(week > (today() - weeks(4)))
  } else{
    df
  }
}

lm_all       = fit(recruitment_workflow,             weekly_recruitment)
lm_thisyear  = fit(recruitment_workflow, filter_sets(weekly_recruitment, "thisyear"))
lm_fourweeks = fit(recruitment_workflow, filter_sets(weekly_recruitment, "fourweeks"))

pred_all       = augment(lm_all,                   study_dates)               %>% mutate(pred_length = "Whole study")
total_all = max(pull(pred_all, .pred)) %>% floor()
pred_thisyear  = augment(lm_thisyear,  filter_sets(study_dates, "thisyear"))  %>% mutate(pred_length = "From Jan'25")
total_thisyear = max(pull(pred_thisyear, .pred)) %>% floor()
pred_fourweeks = augment(lm_fourweeks, filter_sets(study_dates, "fourweeks")) %>% mutate(pred_length = "Last 4 weeks")
total_fourweeks = max(pull(pred_fourweeks, .pred)) %>% floor()

predictions = bind_rows(pred_all, pred_thisyear) %>% 
  bind_rows(pred_fourweeks) %>% 
  mutate(total = floor(.pred)) %>% 
  mutate(pred_length = fct_relevel(pred_length, "Whole study"))
```

::::{.cr-section}

Recruitment commenced in August 2024 and will continue until the end of 2025. @cr-plot1

:::{#cr-plot1}

```{r}
plot1 = weekly_recruitment %>% 
  ggplot(aes(week, total)) +
  geom_line(data = predictions, aes(colour = pred_length, alpha = pred_length), linewidth = 2) +
  geom_point() +
  scale_x_date(breaks = "2 months", date_labels = "%b'%y", expand = c(0.1, 0.1)) +
  scale_y_continuous(expand = c(0.1, 0.1)) +
  geom_label(data = slice_max(weekly_recruitment, total), aes(label = total), size = 7) +
  # geom_label(data = predictions %>% 
  #              group_by(pred_length) %>% 
  #              slice_max(total) %>% 
  #              filter(pred_length == "Whole study"),
  #            aes(label = total), size = 7)   +
  scale_color_manual("Scenario", values = emus_colours) +
  scale_alpha_manual(values = c(0, 0, 0), guide = "none") +
  guides(color = guide_legend(reverse=TRUE)) +
  theme(legend.position = "none",
        legend.position.inside = c(0.1, 0.8),
        plot.margin = margin(0.2, 0.8, 0.2, 0.2, "cm")
  ) +
  labs(y = "Number of patients",
       x = "")

plot1
```

:::

If we recruited with the average rate that we've had since the beginning of the study, we would finish the study with just `r total_all` patients.  @cr-plot2

:::{#cr-plot2}

```{r}

arrows <- 
  tibble(
    x1 = c(ymd("2025-07-05")),
    x2 = c(ymd("2025-11-25")),
    y1 = 800, 
    y2 = 500
  )


plot2 = weekly_recruitment %>% 
  ggplot(aes(week, total)) +
  geom_line(data = predictions, aes(colour = pred_length, alpha = pred_length), linewidth = 2) +
  geom_point() +
  scale_x_date(breaks = "2 months", date_labels = "%b'%y", expand = c(0.1, 0.1)) +
  scale_y_continuous(expand = c(0.1, 0.1)) +
  geom_label(data = slice_max(weekly_recruitment, total), aes(label = total), size = 7) +
  geom_label(data = predictions %>% 
               group_by(pred_length) %>% 
               slice_max(total) %>% 
               filter(pred_length == "Whole study"),
             aes(label = total), size = 7)   +
  scale_color_manual("Scenario", values = emus_colours) +
  scale_alpha_manual(values = c(0.8, 0, 0), guide = "none") +
  guides(color = guide_legend(reverse=TRUE)) +
  theme(legend.position = "none",
        legend.position.inside = c(0.1, 0.8),
        plot.margin = margin(0.2, 0.8, 0.2, 0.2, "cm")
  ) +
  labs(y = "Number of patients",
       x = "") +
  annotate("text", x = ymd("2025-06-01"), y = 875, 
           label = "Number of patients based on \n all time recruitment rate") +
  geom_curve(
    data = arrows, aes(x = x1, y = y1, xend = x2, yend = y2),
    arrow = arrow(length = unit(0.08, "inch")), linewidth = 0.5,
    color = "gray20", curvature = 0.3)

plot2

```
:::


However, as more sites are being opened, our recruitment rate is accelerating.  @cr-plot3

:::{#cr-plot3}

```{r}

arrows <- 
  tibble(
    x1 = c(ymd("2025-07-05")),
    x2 = c(ymd("2025-11-25")),
    y1 = 1000, 
    y2 = 900
  )


plot3 = weekly_recruitment %>% 
  ggplot(aes(week, total)) +
  geom_line(data = predictions, aes(colour = pred_length, alpha = pred_length), linewidth = 2) +
  geom_point() +
  scale_x_date(breaks = "2 months", date_labels = "%b'%y", expand = c(0.1, 0.1)) +
  scale_y_continuous(expand = c(0.1, 0.1)) +
  geom_label(data = slice_max(weekly_recruitment, total), aes(label = total), size = 7) +
  geom_label(data = predictions %>% 
               group_by(pred_length) %>% 
               slice_max(total) %>% 
               filter(pred_length != "Last 4 weeks"),
             aes(label = total), size = 7)   +
  scale_color_manual("Scenario", values = emus_colours) +
  scale_alpha_manual(values = c(0.1, 0.9, 0), guide = "none") +
  guides(color = guide_legend(reverse=TRUE)) +
  theme(legend.position = "none",
        legend.position.inside = c(0.1, 0.8),
        plot.margin = margin(0.2, 0.8, 0.2, 0.2, "cm")
  ) +
  labs(y = "Number of patients",
       x = "") +
  annotate("text", x = ymd("2025-06-01"), y = 1100, 
           label = "Number of patients based on \n Jan-Mar 2025 recruitment rate") +
  geom_curve(
    data = arrows, aes(x = x1, y = y1, xend = x2, yend = y2),
    arrow = arrow(length = unit(0.08, "inch")), linewidth = 0.5,
    color = "gray20", curvature = 0.3)

plot3

```
:::

If we continue recruiting patients with the same rate we have over the last four weeks, we will finish the study with `r total_fourweeks` patients.  @cr-plot4

:::{#cr-plot4}

```{r}

arrows <- 
  tibble(
    x1 = c(ymd("2025-07-05")),
    x2 = c(ymd("2025-11-20")),
    y1 = 1200, 
    y2 = 1250
  )


plot4 = weekly_recruitment %>% 
  ggplot(aes(week, total)) +
  geom_line(data = predictions, aes(colour = pred_length, alpha = pred_length), linewidth = 2) +
  geom_point() +
  scale_x_date(breaks = "2 months", date_labels = "%b'%y", expand = c(0.1, 0.1)) +
  scale_y_continuous(expand = c(0.1, 0.1)) +
  geom_label(data = slice_max(weekly_recruitment, total), aes(label = total), size = 7) +
  geom_label(data = predictions %>% 
               group_by(pred_length) %>% 
               slice_max(total),
             aes(label = total), size = 7)   +
  scale_color_manual("Scenario", values = emus_colours) +
  scale_alpha_manual(values = c(0.1, 0.1, 0.9), guide = "none") +
  guides(color = guide_legend(reverse=TRUE)) +
  theme(legend.position = "none",
        legend.position.inside = c(0.1, 0.8),
        plot.margin = margin(0.2, 0.8, 0.2, 0.2, "cm")
  ) +
  labs(y = "Number of patients",
       x = "") +
  annotate("text", x = ymd("2025-04-01"), y = 1200, 
           label = "Number of patients based on \n last four weeks' recruitment rate") +
  geom_curve(
    data = arrows, aes(x = x1, y = y1, xend = x2, yend = y2),
    arrow = arrow(length = unit(0.08, "inch")), linewidth = 0.5,
    color = "gray20", curvature = 0.1)

plot4

```
:::


Our three recruitment scenarios based on the different rates. @cr-plot5

:::{#cr-plot5}


```{r}
plot_final = weekly_recruitment %>% 
  ggplot(aes(week, total)) +
  geom_line(data = predictions, aes(colour = pred_length, alpha = pred_length), linewidth = 2) +
  geom_point() +
  scale_x_date(breaks = "2 months", date_labels = "%b'%y", expand = c(0.1, 0.1)) +
  scale_y_continuous(expand = c(0.1, 0.1)) +
  geom_label(data = slice_max(weekly_recruitment, total), aes(label = total), size = 7) +
  geom_label(data = predictions %>% 
               group_by(pred_length) %>% 
               slice_max(total),
             aes(label = total), size = 7)   +
  scale_color_manual("Scenario", values = emus_colours) +
  scale_alpha_manual(values = c(0.8, 0.9, 0.9), guide = "none") +
  guides(color = guide_legend(reverse=TRUE)) +
  theme(legend.position = "inside",
        legend.position.inside = c(0.2, 0.8),
        plot.margin = margin(0.2, 0.8, 0.2, 0.2, "cm")
  ) +
  labs(y = "Number of patients",
       x = "")


plot_final
```



:::
